Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-29 paper/paper.tex
pass on section 4.2

The seed rule produced in the first step satisfies the criterion to
only capture positive traffic. However, that rule is overconstrained.
It could include options that are irrelevant to capturing the attack
and could lead to false negatives---the \nids\ may be unable to
capture manifestations of the same attack whose messages do not
exhibit the data satisfying irrelevant options.

@@

\Mar{--------------------------------------- estou aqui}

Note that no rules are produced in the last iteration of the
algorithm. Minimal rules are produced in the second-to-last
iteration. In the other iterations, intermediate rules are
produced. Conceptually, the search induces a search tree with the seed
rule $sr$ as the root, minimal rules as child nodes, and intermediate
rules as internal nodes. The edges of the tree denote the ancestor
relation between two rules. Note also that the algorithm reports every
rule in the tree instead of only minimal rules. The reason is that the
correct rule may be non-minimal. In summary, this algorithm reports
many rules that are plausible by using benign traffic to minimize an
overconstrained rule that is focused on the malicious traffic. In the
following step, \tname{} ranks the rules this algorithm reports by
their similarity with previous rules.

\begin{theorem}
  The minimization algorithm computes rules that capture only
  the positive traffic.
\end{theorem}
\begin{proof}
  Every rule created during the search captures the positive traffic
  as every rule contains a proper subset of the option from the
  original rule and all those options are satisfied by the positive
  traffic. By definition of function $\mathit{isPlausible}$, these rules
  also capture \emph{only} the positive traffic. If some negative
  traffic was captured, a rule would not be added to $\mathcal
  R'$.
\end{proof}

\begin{proposition}
  The minimization algorithm may not produce minimal rules.
\end{proposition}

\begin{proof}
The output of the minimization algorithm may contain rules that are
not minimal given the non-determinism of the algorithm. In principle,
it is possible that options that could be discarded are not selected
in the loop from lines~\ref{start-for}--\ref{end-for}. However, we
found empirically that a small value of $N$ suffices to discard those
options from a rule.
\end{proof}

@@

\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~updating Suricata to only monitor the rule
that is passed as parameter, 2)~replaying the negative traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. It is worth
noting that the benign traffic is stored in \emph{pcap}
files~\cite{pcap} and Suricata can process those files very
efficiently compared to analyzing network messages directly. Instead
of sending traffic to a machine with a deployed server (\eg{}, web
server) that \suri\ monitors, our setup calls Suricata directly
through an API. Suricata is invoked multiple times in the process of
minimizing the rule.  \textbf{Limitations.} It is worth noting that
although \tname{} is accurate for generating rules that discrimate
positive and negative traffic, the quality of these rules is limited
by the quality of the training data it uses. For the positive traffic,
\tname{} only requires one sample, but can benefit of several
samples. The same applies for the negative traffic. In this case,
however, \tname{} can use public data sets available online.

@@

@@

\vspace{1ex}
\subsubsection{Benign traffic.}~Public data sets of bening traffic
exist and can be used to evaluate the rate of false positives of
security tools. We used two popular public sources. The Bigflows.pcap
data set is part of the TcpReplay~\cite{tcpreplay} open source
project. It includes real network traffic of a busy private network
access point to the Internet. We also used data sets from the
Stratosphere IPS Project~\cite{stratosphere-normal} including, for
example, the regular traffic of a university network.

>>>>

\Fix{consider changing the name of this step}
The seed rule produced in the first step satisfies the criterion to
only capture positive traffic. However, that rule is overconstrained.
It could include options that are irrelevant to capturing the attack
and could lead to false negatives---the \nids\ may be unable to
capture manifestations of the same attack whose messages do not
exhibit the data satisfying irrelevant options.

@@

Note that the algorithm produces no rules in the last iteration. Rules
with a single option are produced in the second-to-last iteration. In
other iterations, ules with more than one option are
produced. Conceptually, the search induces a search tree with the seed
rule $sr$ as the root, single-option rules as child nodes, and rules
with more options as internal nodes. The edges of the tree denote the
ancestor relation between two rules.

It is also worth noting that the algorithm reports every rule in the
tree instead of only those that are surely minimal, \ie{} rules with
one option. The reason is that the the traffic used to check
plausibility is limited by definition---reporting only minimal rules
with respect to that traffic could discard the correct rule. In
summary, this algorithm reports many rules that are plausible by using
the benign traffic to reduce the overconstrained rule produced at the
\reveng\ step. In the following step, \tname{} ranks the rules this
algorithm reports by their similarity to existing trusted
rules. Intuitively, rules that do not look alike to rules written by
humans and previously used rank low.

\begin{proposition}
  The minimization algorithm reports rules that capture only the
  positive traffic.
\end{proposition}
\begin{proof}
  Every rule created during the search captures the positive malicious
  traffic as every rule contains a proper subset of the options from
  the original rule and all those options are satisfied by the
  malicious traffic. By definition of function $\mathit{isPlausible}$,
  these rules also capture \emph{only} the positive traffic. If
  negative traffic is captured, a rule is not added to $\mathcal R'$;
  consequently, it cannot be an element of $\mathcal O$ and reported
  on the output.
\end{proof}

%% \begin{proposition}
%%   The minimization algorithm may not produce minimal rules.
%% \end{proposition}

%% \begin{proof}
%% The output of the minimization algorithm may contain rules that are
%% not minimal given the non-determinism of the algorithm. In principle,
%% it is possible that options that could be discarded are not selected
%% in the loop from lines~\ref{start-for}--\ref{end-for}. However, we
%% found empirically that a small value of $N$ suffices to discard those
%% options from a rule.
%% \end{proof}

@@

\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~updating Suricata to only monitor the rule
that is passed as parameter\Mar{isto realmente eh necessario? se sim,
  explique. me pareceu que o que vc. precisa ate aqui jah estaria no
  log do \suri}, 2)~replaying the negative benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. It is worth
noting that the benign traffic is stored in \emph{pcap}
files~\cite{pcap} and Suricata can process those files very
efficiently compared to analyzing network messages directly. Instead
of sending traffic to a machine with a deployed server (\eg{}, web
server) that \suri\ would monitor, our setup calls \suri\ directly
through an API. \suri\ is invoked multiple times in the process of
minimizing the rule.  \textbf{Limitations.} It is worth noting that
although \tname{} is accurate for generating rules that discrimate
positive and negative traffic, the quality of these rules is limited
by the quality of the training data it uses. For the positive traffic,
\tname{} only requires one sample, but can benefit of several
samples\Mar{como? nao estah claro. por favor expliquem.}. The same
applies to the negative traffic. However, in this case, \tname{} uses
different public data sets available online (see
Section~\ref{sec:objectanalysis-benign}).

@@

\Mar{----------------------------- estou aqui}

@@

\vspace{1ex}
\subsubsection{\label{sec:objectanalysis-benign}Benign traffic.}~Public data sets of bening traffic
exist and can be used to evaluate the rate of false positives of
security tools. We used two popular public sources. The Bigflows.pcap
data set is part of the TcpReplay~\cite{tcpreplay} open source
project. It includes real network traffic of a busy private network
access point to the Internet. We also used data sets from the
Stratosphere IPS Project~\cite{stratosphere-normal} including, for
example, the regular traffic of a university network.
